---
title: "Gamma Likelihood Plot"
author: "Jonathan Luu"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Method
The gamma log likelihood equation is defined by the function loglik.gamma. For this problem, alpha is known and set to 3, without loss of generality. As beta is unknown, a sequence of beta values (length=19000) is generated and used to estimate the maximum likelihood value.  

```{r}
set.seed(123)
alpha <- 3
n <- 2000

# Gamma log likelihood equation
# Alpha is known = 3
# Beta is the unknown parameter
loglik.gamma <- function(beta, alpha, x){
  logl<-sum(dgamma(x,alpha,scale=beta,log=TRUE))
  return(logl)
}

runAndPlot<- function(x){
  # Experimental iteration
  beta_experiment<-seq(1,20, by=0.001)
  beta_mle<-sapply(beta_experiment, loglik.gamma, alpha=alpha, x=x)
  paired<-data.frame(beta_experiment, beta_mle)
  ans<-paired[which.max(paired$beta_mle),]
  
  # Plot the paired beta and mle values
  plot(beta_experiment,beta_mle,type='l',xlab=expression(beta),
       ylab=expression(L( ~ beta  ,x)), main="Beta_hat x Log likelihood for Gamma Distribution")
  abline(v=ans$beta_experiment,col="red",lty=2)
  abline(h=ans$beta_mle,col="red",lty=2)
  
  print(ans)
}
```

\newpage
For the first dataset, the following parameters were used: n=2000, alpha=3, beta=3
```{r}
# generate gamma(3,3) data
x <- rgamma(n, alpha, scale=3)
runAndPlot(x)
```

The estimated value of beta is 2.943, which is very close to the actual beta value of 3.

\newpage
For the second dataset, the following parameters were used: n=2000, alpha=3, beta=12
```{r}
# generate gamma(3,12) data
y <- rgamma(n, alpha, scale=12)
runAndPlot(y)
```

The estimated value of beta is 11.922, which is very close to the actual beta value of 12.