---
title: "PM522B HW1"
author: "Jonathan Luu"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 3
For X1, X2, ..., Xn ~ Unif[0, 1], write a function in R to draw random samples of size 15, and take the maximum of each sample.

```{r}
#Set seed and global variables
set.seed(120)
NumberOfSims<-10000
NumberOfSamples<-15

#Storage container for maximum values
StorageVector<-double(NumberOfSims)
for (i in 1:NumberOfSims){
  #Draw 15 random samples, take the maximum, and store it
  StorageVector[i]=max(runif(NumberOfSamples,0 ,1))
}

#Graph the distribution
hist(StorageVector, xlab="Maximum values from each sample", 
     main="Histogram of maximum values from each sample")
```

## Question 4
Using kids.RData posted on Blackboard (this is the CHS data used in PM511a), construct Q-Q plots in R comparing the wt variable (weight in lbs) to the normal, log-normal, gamma, and Weibull distributions. Do not use a package to create these plots. Label all axes accordingly. To what distribution do the data follow most closely?

```{r}
load("C:\\Users\\jluu\\Dropbox\\School\\PM522B\\HW1\\kids.RData")

#Get kids dataset quantiles
kids_quantiles<-quantile(kids$wt, probs=seq(0,1,1/length(kids$wt)))

#Generate normal quantile distribution
normal_quantiles<-qnorm(seq(0,1,1/length(kids$wt)), mean(kids$wt), sd(kids$wt))
plot(normal_quantiles, kids_quantiles, xlab="Normal quantiles", ylab="Sample quantiles",
     main="Normal QQ plot")
abline(0,1)
```

\newpage
```{r}
#Generate lognormal quantile distribution
lognormal_quantiles<-qlnorm(seq(0,1,1/length(kids$wt)), mean(log(kids$wt)), sd(log(kids$wt)))
plot(lognormal_quantiles, kids_quantiles, xlab="Lognormal quantiles", ylab="Sample quantiles",
     main="Lognormal QQ plot")
abline(0,1)
```

\newpage
```{r}
#Generate gamma quantile distribution
alpha=(mean(kids$wt)^2)/var(kids$wt)
beta=var(kids$wt)/mean(kids$wt)
gamma_quantiles<-qgamma(seq(0,1,1/length(kids$wt)), alpha, beta)
plot(gamma_quantiles, kids_quantiles, xlab="Gamma quantiles", ylab="Sample quantiles",
      main="Gamma QQ plot")
abline(0,1)
```

\newpage
```{r}
#Generate weibull quantile distribution
weibull_quantiles<-qweibull(seq(0,1,1/length(kids$wt)), alpha, beta)
plot(weibull_quantiles, kids_quantiles, xlab="Weibull quantiles", ylab="Sample quantiles",
      main="Weibull QQ plot")
abline(0,1)
```

Out of these four QQ plots, the lognormal QQ plot fits the best.

